Notes: New gaugings in 3d from three-form

### 2023-06-20 Draft_3dH3.tex
Moved section 4.2 to an appendix, changed some typo in the text, added Berg and Samtleben in the conclusions.

Nonetheless, there are half-maximal, \textit{i.e.} ${\cal N}=(4,4)$, supersymmetric AdS$_3$ vacua of string theory that are extremely intriguing. These are the AdS$_3 \times S^3 \times S^3 \times S^1$ and AdS$_3 \times S^3 \times T^4$ of IIB string theory (realising the ``large'' and ``small'' ${\cal N}=(4,4)$ superconformal symmetries), and, unlike in higher dimensions, can be supported by pure NS-NS flux. Not only does this mean that there are also heterotic versions of these vacua (preserving ${\cal N}=(4,0)$ supersymmetry), but also that they can be readily studied via the string worldsheet CFT \cite{Maldacena:2000hw,Eberhardt:2018ouy,Eberhardt:2019niq,Eberhardt:2019ywk}. \EM{Cut?} Moreover, the string sigma model in these vacua is also integrable \cite{}. Finally, the AdS$_3$/CFT$_2$ correspondence has seen much progress recently, with many new holographic duals being established \cite{}.

@@

Another technical result of our work is the derivation of the potential of 3-dimensional gauged supergravity \cite{Nicolai:2000sc,Nicolai:2001sv} in terms of the embedding tensor. The potential was previously only known in terms of fermion shift matrices. As a by-product, we find a constraint that must be obeyed by the embedding tensor of any 3-dimensional gauged supergravity that can be uplifted to 10-/11-dimensional supergravity. As we show, this allows to prove the lack of a higher-dimensional origin of some 3-dimensional gaugings, as we illustrate on an example. \EM{update}

The outline of our paper is as follows. We begin with a review of $\EE$ ExFT in section \ref{s:Review} and how to derive consistent truncations in this formalism in \ref{s:TruncationReview}. In section \ref{s:Potential} we derive the potential of 3-dimensional gauged supergravity in terms of the embedding tensor and prove the lack of higher-dimensional origin of some 3-dimensional theories. Then, we show how to add fluxes to consistent truncations in \ref{s:AddFlux}, allowing us to construct the consistent truncation of IIB supergravity on $S^3 \times S^3 \times S^1$. Using this truncation, we study the moduli space of the AdS$_3$ vacua in section \ref{s:Moduli}, before concluding in section \ref{s:Conclusions}.

@@

\section{Deriving the potential of ${\cal N}=16$ gauged supergravity} \label{s:Potential}
Here we will use the $\EE$ ExFT Lagrangian and the consistent truncation Ansatz \eqref{eq:GSSansatz} to derive the potential of three-dimensional ${\cal N}=16$ gauged supergravity \cite{Nicolai:2000sc,Nicolai:2001sv} in terms of the embedding tensor. Not only is this an interesting application of ExFT, relying on purely bosonic considerations and bypassing the usual construction of the gauged supergravity potential using supersymmetry\footnote{Note that the same strategy was recently used to impressively derive the potential of maximal two-dimensional gauged supergravity \cite{Bossard:2022wvi}, where fermions are extremely poorly understood. As a result, in that case, the bosonic $\En{9}$ ExFT \cite{Bossard:2018utw,Bossard:2021jix} provides the only currently accessible route for computing the gauged supergravity potential.}, but the potential is crucial for finding vacua of the three-dimensional theory and uplifting these in the later parts of this paper. In addition, we derive a necessary condition for the embedding tensor to arise from a compactification of higher-dimensional supergravity and illustrate it on an example.

\subsection{Expression of the potential in terms of the embedding tensor}
In order to derive the potential, we adopt the following strategy. We want $\mathscr{L}_{\rm int}$ \eqref{eq:ExFTLint} to reduce to the embedding tensor squared upon inserting the generalised Scherk-Schwarz Ans\"{a}tze \eqref{eq:GSSansatz}. However, as in higher dimensions, \textit{e.g.} \cite{Berman:2012uy,Musaev:2013rq,Blair:2014zba}, this does not have to match identically, but only up to total derivative terms and terms which violate the section condition. The possible boundary terms are
\begin{equation} \label{eq:BoundaryTerms}
	\frac{1}{\sqrt{\dg}} \partial_M \left(\sqrt{\dg} \partial_N \gM^{MN} \right) \quad \text{and} \quad \frac{1}{\sqrt{\dg}} \partial_M \left( \gM^{MN} \partial_N \sqrt{\dg}\right) \,.
\end{equation}
However, since the ${\cal N}=16$ supergravity potential will be given by terms quadratic in $X_{\fl{MN}}{}^{\fl{P}}$, we do not want the boundary terms to involve any double derivative terms. Therefore, the two boundary terms \eqref{eq:BoundaryTerms} can only appear via the combination
\begin{equation} \label{eq:BoundaryTermCombination}
	\frac{1}{\sqrt{\dg}}\,\partial_{M}\Big(\sqrt{\dg}\,\partial_{N}\M^{MN}+\frac{4}{3}\,\M^{MN}\partial_{N}\sqrt{\dg}\Big) \,.
\end{equation}

@@

It turns out that for gaugings that arise from consistent truncations, these two terms \eqref{eq:ExtraX2Terms} can be related to the square of the trombone tensor. We can square eq.~\eqref{eq:symembeddingtensor} to derive an expression for $X_{\ov{MN}}X^{\ov{MN}}$ in terms of the current $\Gamma$. We find
\begin{equation}\label{eq:XXwitheta}
	X_{\ov{MN}}X^{\ov{MN}} = 21 \rho^{-2}\, \Gamma_{\ov{MN}}\, \Gamma^{\ov{NM}} + 19 \rho^{-2}\left(\Gamma_{\ov M}{}^{\ov M}\right)^2 - 15\, \xi_{\ov{M}}\, \xi^{\ov{M}} \,.
\end{equation}
Similarly, squaring the trombone tensor $\xi$ gives
\begin{equation}\label{eq:xixiwitheta}
	\xi_{\ov M}\,\xi^{\ov M}=\rho^{-2}\left(\Gamma_{\ov M}{}^{\ov M}\Gamma_{\ov N}{}^{\ov N}+\Gamma_{\ov{MN}}\Gamma^{\ov{NM}}\right) \,.
\end{equation}
We can now deduce
\begin{equation}
	X_{\ov{MN}} \,X^{\ov{MN}}= 6\, \xi_{\ov M} \, \xi^{\ov M} - 2\, \rho^{-2}\left(\Gamma_{\ov M}{}^{\ov M}\right)^2 \,,
\end{equation}
which can be expressed entirely in terms of $X_{\fl{M}\fl{N}}$ after tracing eq.~\eqref{eq:symembeddingtensor}:
\begin{equation}\label{eq:reltrX}
	X_{\ov{MN}}\,X^{\ov{MN}} = 6\, \xi_{\ov M}\, \xi^{\ov M} - \frac1{1922}\left(X_{\ov M}{}^{\ov M}\right)^2 \,. 
\end{equation}
Thus, \eqref{eq:reltrX} is a constraint on the embedding tensor of three-dimensional supergravity which must be satisfied in order for it to have an uplift to 10-/11-dimensional supergravity. Note that this constraint is not implied by the quadratic constraint. To see this, simply observe that the theory with $\En{8}$ gauging has $X_{\ov{MN}}=\eta_{\ov{MN}}$ and $\xi_{\fl{M}} = 0$ \cite{Nicolai:2000sc,Nicolai:2001sv}, violating \eqref{eq:reltrX}.

@@

\subsection{Ruling out gaugings}
\EM{Let's check the $\GL{6} \times \SO{2,2}$ gauging (half-maximal embedded in maximal)....}

As mentioned above, eq.~\eqref{eq:reltrX} can be used to assess whether a given three-dimensional theory of embedding tensor $X_{\fl{MN}}$ has a higher-dimensional origin. Let us illustrate that on the ${\cal N}=(8,0)$ ${\rm AdS}_{3}$ vacuum with $\GL{6} \times \SO{2,2}$ gauging constructed in ref.~\cite{Deger:2019tem}. This solution has a compact gauging smaller than $\SO{9}$, thus escaping the no-go result of ref.~\cite{Galli:2022idq}. The embedding tensor of the solution is expressed by breaking $\EE$ as
\begin{equation}
	\begin{aligned}
		\EE &\longrightarrow \SO{8,8}\\
		\mathbf{248} &\longrightarrow \mathbf{120} \oplus \mathbf{128}\\
		X^{\fl{M}} &\longrightarrow \left\{X^{[\fl{\sf IJ}]},X^{\fl{\sf A}}\right\},
	\end{aligned}
\end{equation}
with $\fl{\sf I}$ and $\fl{\sf A}$ labels of the fundamental and spinorial representations of $\SO{8,8}$, respectively. Explicitly,
\begin{equation}
	\begin{aligned}
		X_{\fl{\sf IJ},\fl{\sf KL}} &= \Theta_{\fl{\sf IJ},\fl{\sf KL}},\\
		X_{\fl{\sf AB}} &= -2\,\theta\,\eta_{\fl{\sf AB}}+\frac{1}{24}\,\Gamma^{\fl{\sf IJ},\fl{\sf KL}}_{\fl{\sf AB}}\,\Theta_{\fl{\sf IJKL}},\\
		X_{\fl{\sf IJ},\fl{\sf A}} &=0,
	\end{aligned}
\end{equation}
where $\Gamma^{\fl{\sf IJKL}}_{\fl{\sf AB}}$ is the four-fold product of $\SO{8,8}$ $\Gamma$ matrices and $\eta_{\fl{\sf AB}}$ the $\mathbf{128}\otimes\mathbf{128}$ component of the Cartan-Killing metric. The tensor $\Theta_{\fl{\sf IJ},\fl{\sf KL}}$ is given by
\begin{equation}
	\Theta_{\fl{\sf IJ},\fl{\sf KL}}=\theta_{\fl{\sf IJKL}}+\frac12\big(\eta_{\fl{\sf K}[\fl{\sf I}}\theta_{\fl{\sf J}]\fl{\sf L}}-\eta_{\fl{\sf L}[\fl{\sf I}}\theta_{\fl{\sf J}]\fl{\sf K}}\big)+\theta\,\eta_{\fl{\sf K}[\fl{\sf I}}\eta_{\fl{\sf J}]\fl{\sf L}}\,,
\end{equation}
with $\theta_{\fl{\sf IJKL}}$ totally antisymmetric and $\theta_{\fl{\sf IJ}}$ symmetric, with non vanishing components
\begin{align}
  & \theta_{ij}=4\,\delta_{ij}\;,\quad \theta_{\alpha\beta}=-12\,\delta_{\alpha\beta}\;,\quad \theta_{r_{+}s_{+}}=-4\,\delta_{r_{+}s_{+}}\;,\quad \theta_{r_{-}s_{-}}=12\,\delta_{r_{-}s_{-}}\;,\nonumber\\
  & \theta=-8\;,\quad \theta_{\alpha\beta r_{-}s_{-}} = -48\,\sigma_{\alpha\beta}\,\sigma_{r_{-}s_{-}}\;,\quad \theta_{ij r_{+}s_{+}} = 32\,\delta_{i[r_{+}}\,\delta_{s_{+}]j}\;,
\end{align}
where... \CE{(explain all notations...)} With this embedding, we get
\begin{equation}
	X_{\ov{MN}}\,X^{\ov{MN}} = \frac{55}{1922}\left(X_{\ov M}{}^{\ov M}\right)^2,
\end{equation}
thus violating the condition~\eqref{eq:reltrX} (the trombone gauging vanishes for the solution). Then, we deduce that the ${\cal N}=(8,0)$ solution of ref.~\cite{Deger:2019tem} with gauging $\GL{6} \times \SO{2,2}$ does not admit a higher-dimensional origin. This illustrate how the constraint~\eqref{eq:reltrX} can be used to determine if a given three-dimensional vacuum can be uplifted.

@@

This motivates the following question: Given a consistent truncation, \textit{i.e.} $\cU_{\fl{M}}{}^M$, satisfying \eqref{eq:GenLeibniz}, when can we add a new flux component of string theory to the compactification to obtain a new consistent truncation? Adding a new flux component to the truncation is equivalent to twisting the generalised frame as follows \EM{Email DW and MP to ask who we should cite for this?}
\begin{equation} \label{eq:Twisting}
	\cU_{\fl{M}}{}^M \longrightarrow \cU'_{\fl{M}}{}^M = \cU_{\fl{M}}{}^N \exp(C)_N{}^M \,,
\end{equation}
where $C$ denotes the $\EE$ generator corresponding to the potential we want to add to the compactification. The effect of the twist \eqref{eq:Twisting} is that the generalised Lie derivative of $\cU'$ satisfies
\begin{equation}
	\gL_{\ff{U}'_{\fl{M}}} \cU'_{\fl{N}}{}^M = \left( \gL_{\ff{U}_{\fl{M}}} \cU_{\fl{N}}{}^N + F_{PQ}{}^N\, \cU_{\fl{M}}{}^P\, \cU_{\fl{N}}{}^Q \right) \exp(C)_N{}^M \,,
\end{equation}
where $F_{MN}{}^P$ is a tensor in the $\mathbf{1} \oplus \mathbf{248} \oplus \mathbf{3875}$ of $\EE$, \textit{i.e.} the same representation as the embedding tensor, corresponding to the field strength of the potential $C$ in \eqref{eq:Twisting}. Using \eqref{eq:GenLeibniz}, we now have
\begin{equation}
	\gL_{\ff{U}'_{\fl{M}}} \cU'_{\fl{N}}{}^M = \left( X_{\fl{M}\fl{N}}{}^{\fl{P}} + \rho^{-1}\, F_{\fl{M}\fl{N}}{}^{\fl{P}} \right) \cU'_{\fl{P}}{}^M \,,
\end{equation}
where we defined
\begin{equation} \label{eq:FlatF}
	F_{\fl{M}\fl{N}}{}^{\fl{P}} \equiv \UI_{\fl{M}}{}^M\, \UI_{\fl{N}}{}^N\, U_P{}^{\fl{P}}\, F_{MN}{}^P \,,
\end{equation}
and $X_{\fl{M}\fl{N}}{}^{\fl{P}}$ is already constant. Therefore, we have a consistent truncation if and only if $\rho^{-1}\, F_{\fl{M}\fl{N}}{}^{\fl{P}}$ is constant.

@@

\subsubsection{The $S^n$ twist matrix} \label{s:SLn}
\EM{Move to appendix?!?!} \CE{(I think I would say yes. Work here.)}\\
Since we will use it throughout this paper, let us construct the $\SL{4} \simeq \SO{3,3}$ twist matrix that describes the consistent truncation on $S^3$. This can be viewed as a special case of a family of consistent truncations on $S^n$ \cite{Lee:2014mla,Hohm:2014qga,Baguet:2015iou}. Here, we give a slightly different, but equivalent, form of this construction.

@@

We denote by $U_A{}^{\ov{A}}$ the $\SO{3,3}$ twist matrix corresponding to the $S^3$ truncation constructed in \cite{Lee:2014mla,Baguet:2015iou} and obtained from \eqref{eq:SLnFrame}, \eqref{eq:rho}, \eqref{eq:SLnTwist}. Then, we can add a 3-form potential by twisting with the $\SO{4,4}$ generator \CE{(Sure here? Check notes.)}
\begin{equation} \label{eq:C2S3example}
	(e^C)^{MN} = C_A \left( t^A \right)^{MN} \,,
\end{equation}
with $C_A$ an element of the $\mathbf{6_{2}}$, and $\left( t^A \right)^{MN}$ the generator corresponding to $\SO{4,4}$ whose only non-zero component is
\begin{equation}
	\left(t^A\right)^{Bz} = \eta^{AB} \,.
\end{equation}
Here $\eta_{MN}$ and $\eta_{AB}$ are the $\SO{4,4}$ and $\SO{3,3}$ invariant metrics, respectively, and are used to raise/lower the corresponding vector indices. Decomposing with respect to the geometric $\SL{3} \times \mathbb{R}^+$ of the $S^3$,  the coordinates on $S^3$ live in the $Y^A = \left( y^i,\, y_i \right)$, $i = 1, 2, 3$, while $C_A = \left( C_i,\, C^i \right)$ naturally contains a 2-form $C^i$. The field strength $H_3 = \partial^A C_A = \partial_i C^i$ is a singlet of $\SO{3,3}$ and thus we see that $\rho^{-1} F_{\ov{M}\ov{N}}{}^{\ov{P}}$ in \eqref{eq:FlatF} is constant and we obtain a consistent truncation with a new 3-form flux. Evaluating the twist matrix
\begin{equation}
	U'_M{}^{\ov{M}} = (e^C)_M{}^N  U_N{}^{\ov{M}} \,,
\end{equation}
explicitly using \eqref{eq:C2S3example} and setting $C^i = \rho^{-1}\,\lambda\, \xi^i$, with $\nabla_i \xi^i = 1$, in the notation of \cite{Eloy:2021fhc}, we obtain precisely the twist matrix used in \cite{Eloy:2021fhc} for the consistent truncation of 6-dimensional ${\cal N}=(1,1)$ supergravity on $S^3$ with $H_3$ flux. %\EM{Give precise form of $C^i$ such that we recover $\lambda$?}

@@

Finally, the metric parametrising this moduli space is:
\begin{equation}
	\begin{aligned}
		\d s^{2}_{\rm mod.} &= -\frac{1}{2}\bigg[e^{2\psi}\left(e^{-\Omega}\,\d\zeta_{1}^{2}+e^{\Omega}\,\d\zeta_{2}^{2}+e^{\ts{\Omega}}\,\d\ts{\zeta}_{1}^{2}+e^{-\ts{\Omega}}\,\d\ts{\zeta}_{2}^{2}\right) +2\,\d\psi^{2} + \frac{1}{2}\,\d\Omega^{2}+ \frac{1}{2}\,\d\ts{\Omega}^{2}\bigg]\,.
	\end{aligned}
\end{equation}
As previously, the deformation breaks the isometry group down to ${\rm U}(1)_{\rm L}\times{\rm U}(1)_{\rm R}\times\ts{{\rm U}(1)}_{\rm L}\times\ts{{\rm U}(1)}_{\rm R}$ and all supersymmetries. Note that for
\begin{equation}
	\zeta_{1}=\pm\zeta_{2},\quad \ts{\zeta}_{1}=\mp\ts{\zeta}_{2},\quad\Omega=\ts{\Omega}=0,
\end{equation}
the solution~\eqref{eq:metriczeta} reproduces eq.~\eqref{eq:chipsi} for \CE{(Double check.)}
\begin{equation}
	\chi_{1}=\pm\chi_{2}=\frac{\zeta_{1}}{1+\zeta_{1}^{2}+\ts{\zeta}_{1}^{2}},\quad \ts{\chi}_{1}=\mp\ts{\chi}_{2}=-\frac{\ts{\zeta}_{1}}{1+\zeta_{1}^{2}+\ts{\zeta}_{1}^{2}},\quad e^{2\psi}=\frac{1}{(1+\zeta_{1}^{2}+\ts{\zeta}_{1}^{2})^{2}},
\end{equation}
with ${\cal N}=(4,0)$ (or $(0,4)$) and symmetry enhancement to ${\rm SO}(4)\times{\rm U}(1)^{2}$. There seems to be no other points with (super)-symmetry enhancement in the scalar manifold.

@@

\section{Conclusions} \label{s:Conclusions}
In this paper, we used $\EE$ ExFT to construct the consistent truncation of IIB supergravity around the ${\cal N}=(4,4)$ AdS$_3 \times S^3 \times S^3 \times S^1$ vacuum. Even though the vacuum breaks half of the supersymmetries, we are able to construct a consistent truncation to 3-dimensional maximal gauged supergravity, providing an uplift to the ${\cal N}=16$ gauged supergravity proposed in \cite{Hohm:2005ui} for the 3-dimensional description of this AdS$_3$ vacuum. We constructed our consistent truncation by doing a $S^1$ reduction of the truncation of IIB on $S^3 \times S^3$ \cite{Inverso:2016eet}, and turning on an additional $H_7$ flux on $S^3 \times S^3 \times S^1$, which is necessary to obtain an AdS$_3$ vacuum. We also show that it is not possible to turn on such an additional flux for the other ``dyonic'' consistent truncations on $S^6$, $S^5 \times S^1$, $S^4 \times S^2$ \cite{Guarino:2015jca,Inverso:2016eet} reduced on $S^1$, whilst maintining a consistent truncation to ${\cal N}=16$ gauged supergravity.

>>>>

Nonetheless, there are half-maximal, \textit{i.e.} ${\cal N}=(4,4)$, supersymmetric AdS$_3$ vacua of string theory that are extremely intriguing. These are the AdS$_3 \times S^3 \times S^3 \times S^1$ and AdS$_3 \times S^3 \times T^4$ of IIB string theory (realising the ``large'' and ``small'' ${\cal N}=(4,4)$ superconformal symmetries), and, unlike in higher dimensions, can be supported by pure NS-NS flux. Not only does this mean that there are also heterotic versions of these vacua (preserving ${\cal N}=(4,0)$ supersymmetry), but also that they can be readily studied via the string worldsheet CFT \cite{Maldacena:2000hw,Eberhardt:2018ouy,Eberhardt:2019niq,Eberhardt:2019ywk}. %\EM{Cut?} Moreover, the string sigma model in these vacua is also integrable \cite{}. Finally, the AdS$_3$/CFT$_2$ correspondence has seen much progress recently, with many new holographic duals being established \cite{}.

@@

Another technical result of our work is the derivation of the potential of 3-dimensional gauged supergravity \cite{Nicolai:2000sc,Nicolai:2001sv} in terms of the embedding tensor. The potential was previously only known in terms of fermion shift matrices. As a by-product, we find a constraint that must be obeyed by the embedding tensor of any 3-dimensional gauged supergravity that can be uplifted to 10-/11-dimensional supergravity. This allows to prove the lack of a higher-dimensional origin of some 3-dimensional gaugings, as we illustrate on an example. \EM{update}

The outline of our paper is as follows. We begin with a review of $\EE$ ExFT in section \ref{s:Review} and how to derive consistent truncations in this formalism in \ref{s:TruncationReview}. In section \ref{s:Potential} we derive the potential of 3-dimensional gauged supergravity in terms of the embedding tensor and prove the lack of higher-dimensional origin of some 3-dimensional theories. Then, we show how to add fluxes to consistent truncations in \ref{s:AddFlux}, allowing us to construct the consistent truncation of IIB supergravity on $S^3 \times S^3 \times S^1$. Using this truncation, we study the moduli space of the AdS$_3$ vacua in section \ref{s:Moduli}, before concluding in section \ref{s:Conclusions}. We end the paper by an example in appendix~\ref{app:rulingoutgauging} of how to check if a three-dimensional supergravity can be uplifted, and a discussion of the $D^{1}(2,1;\alpha)$ supermultiplets in appendix~\ref{app:d21a}.

@@

\section{Deriving the potential of ${\cal N}=16$ gauged supergravity} \label{s:Potential}
Here we will use the $\EE$ ExFT Lagrangian and the consistent truncation Ansatz \eqref{eq:GSSansatz} to derive the potential of three-dimensional ${\cal N}=16$ gauged supergravity \cite{Nicolai:2000sc,Nicolai:2001sv} in terms of the embedding tensor. Not only is this an interesting application of ExFT, relying on purely bosonic considerations and bypassing the usual construction of the gauged supergravity potential using supersymmetry\footnote{Note that the same strategy was recently used to impressively derive the potential of maximal two-dimensional gauged supergravity \cite{Bossard:2022wvi}, where fermions are extremely poorly understood. As a result, in that case, the bosonic $\En{9}$ ExFT \cite{Bossard:2018utw,Bossard:2021jix} provides the only currently accessible route for computing the gauged supergravity potential.}, but the potential is crucial for finding vacua of the three-dimensional theory and uplifting these in the later parts of this paper.

% In addition, we derive a necessary condition for the embedding tensor to arise from a compactification of higher-dimensional supergravity and illustrate it on an example.
% \subsection{Expression of the potential in terms of the embedding tensor}

In order to derive the potential, we adopt the following strategy. We want $\mathscr{L}_{\rm int}$ \eqref{eq:ExFTLint} to reduce to the embedding tensor squared upon inserting the generalised Scherk-Schwarz Ans\"{a}tze \eqref{eq:GSSansatz}. However, as in higher dimensions, \textit{e.g.} \cite{Berman:2012uy,Musaev:2013rq,Blair:2014zba}, this does not have to match identically, but only up to total derivative terms and terms which violate the section condition. The possible boundary terms are
\begin{equation} \label{eq:BoundaryTerms}
	\frac{1}{\sqrt{\dg}} \partial_M \left(\sqrt{\dg} \partial_N \gM^{MN} \right) \quad \text{and} \quad \frac{1}{\sqrt{\dg}} \partial_M \left( \gM^{MN} \partial_N \sqrt{\dg}\right) \,.
\end{equation}
However, since the ${\cal N}=16$ supergravity potential will be given by terms quadratic in $X_{\fl{MN}}{}^{\fl{P}}$, we do not want the boundary terms to involve any double derivative terms. Therefore, the two boundary terms \eqref{eq:BoundaryTerms} can only appear via the combination
\begin{equation} \label{eq:BoundaryTermCombination}
	\frac{1}{\sqrt{\dg}}\,\partial_{M}\Big(\sqrt{\dg}\,\partial_{N}\M^{MN}+\frac{4}{3}\,\M^{MN}\partial_{N}\sqrt{\dg}\Big) \,.
\end{equation}

@@

It turns out that for gaugings that arise from consistent truncations, these two terms \eqref{eq:ExtraX2Terms} can be related to the square of the trombone tensor. We can square eq.~\eqref{eq:symembeddingtensor} to derive an expression for $X_{\ov{MN}}X^{\ov{MN}}$ in terms of the current $\Gamma$. We find
\begin{equation}\label{eq:XXwitheta}
	X_{\ov{MN}}X^{\ov{MN}} = 21 \rho^{-2}\, \Gamma_{\ov{MN}}\, \Gamma^{\ov{NM}} + 19 \rho^{-2}\left(\Gamma_{\ov M}{}^{\ov M}\right)^2 - 15\, \xi_{\ov{M}}\, \xi^{\ov{M}} \,.
\end{equation}
Similarly, squaring the trombone tensor $\xi$ gives
\begin{equation}\label{eq:xixiwitheta}
	\xi_{\ov M}\,\xi^{\ov M}=\rho^{-2}\left(\Gamma_{\ov M}{}^{\ov M}\Gamma_{\ov N}{}^{\ov N}+\Gamma_{\ov{MN}}\Gamma^{\ov{NM}}\right) \,.
\end{equation}
We can now deduce
\begin{equation}
	X_{\ov{MN}} \,X^{\ov{MN}}= 6\, \xi_{\ov M} \, \xi^{\ov M} - 2\, \rho^{-2}\left(\Gamma_{\ov M}{}^{\ov M}\right)^2 \,,
\end{equation}
which can be expressed entirely in terms of $X_{\fl{M}\fl{N}}$ after tracing eq.~\eqref{eq:symembeddingtensor}:
\begin{equation}\label{eq:reltrX}
	X_{\ov{MN}}\,X^{\ov{MN}} = 6\, \xi_{\ov M}\, \xi^{\ov M} - \frac1{1922}\left(X_{\ov M}{}^{\ov M}\right)^2 \,. 
\end{equation}
Thus, \eqref{eq:reltrX} is a constraint on the embedding tensor of three-dimensional supergravity which must be satisfied in order for it to have an uplift to 10-/11-dimensional supergravity (see appendix~\ref{app:rulingoutgauging} for an example). Note that this constraint is not implied by the quadratic constraint. To see this, simply observe that the theory with $\En{8}$ gauging has $X_{\ov{MN}}=\eta_{\ov{MN}}$ and $\xi_{\fl{M}} = 0$ \cite{Nicolai:2000sc,Nicolai:2001sv}, violating \eqref{eq:reltrX}.

@@

@@

This motivates the following question: Given a consistent truncation, \textit{i.e.} $\cU_{\fl{M}}{}^M$, satisfying \eqref{eq:GenLeibniz}, when can we add a new flux component of string theory to the compactification to obtain a new consistent truncation? Adding a new flux component to the truncation is equivalent to twisting the generalised frame as follows %\EM{Email DW and MP to ask who we should cite for this?}
\begin{equation} \label{eq:Twisting}
	\cU_{\fl{M}}{}^M \longrightarrow \cU'_{\fl{M}}{}^M = \cU_{\fl{M}}{}^N \exp(C)_N{}^M \,,
\end{equation}
where $C$ denotes the $\EE$ generator corresponding to the potential we want to add to the compactification. The effect of the twist \eqref{eq:Twisting} is that the generalised Lie derivative of $\cU'$ satisfies
\begin{equation}
	\gL_{\ff{U}'_{\fl{M}}} \cU'_{\fl{N}}{}^M = \left( \gL_{\ff{U}_{\fl{M}}} \cU_{\fl{N}}{}^N + F_{PQ}{}^N\, \cU_{\fl{M}}{}^P\, \cU_{\fl{N}}{}^Q \right) \exp(C)_N{}^M \,,
\end{equation}
where $F_{MN}{}^P$ is a tensor in the $\mathbf{1} \oplus \mathbf{248} \oplus \mathbf{3875}$ of $\EE$, \textit{i.e.} the same representation as the embedding tensor, corresponding to the field strength of the potential $C$ in \eqref{eq:Twisting}. Using \eqref{eq:GenLeibniz}, we now have
\begin{equation}
	\gL_{\ff{U}'_{\fl{M}}} \cU'_{\fl{N}}{}^M = \left( X_{\fl{M}\fl{N}}{}^{\fl{P}} + \rho^{-1}\, F_{\fl{M}\fl{N}}{}^{\fl{P}} \right) \cU'_{\fl{P}}{}^M \,,
\end{equation}
where we defined
\begin{equation} \label{eq:FlatF}
	F_{\fl{M}\fl{N}}{}^{\fl{P}} \equiv \UI_{\fl{M}}{}^M\, \UI_{\fl{N}}{}^N\, U_P{}^{\fl{P}}\, F_{MN}{}^P \,,
\end{equation}
and $X_{\fl{M}\fl{N}}{}^{\fl{P}}$ is already constant. Therefore, we have a consistent truncation if and only if $\rho^{-1}\, F_{\fl{M}\fl{N}}{}^{\fl{P}}$ is constant.

@@

\subsubsection{The $S^n$ twist matrix} \label{s:SLn}
\EM{Move to appendix?!?!} \CE{(I think I would say yes.)}\\
Since we will use it throughout this paper, let us construct the $\SL{4} \simeq \SO{3,3}$ twist matrix that describes the consistent truncation on $S^3$. This can be viewed as a special case of a family of consistent truncations on $S^n$ \cite{Lee:2014mla,Hohm:2014qga,Baguet:2015iou}. Here, we give a slightly different, but equivalent, form of this construction.

@@

We denote by $U_A{}^{\ov{A}}$ the $\SO{3,3}$ twist matrix corresponding to the $S^3$ truncation constructed in \cite{Lee:2014mla,Baguet:2015iou} and obtained from \eqref{eq:SLnFrame}, \eqref{eq:rho}, \eqref{eq:SLnTwist}. Then, we can add a 3-form potential by twisting with the $\SO{4,4}$ generator \CE{(Sure here?)}
\begin{equation} \label{eq:C2S3example}
	(e^C)^{MN} = \eta^{MN} + C_A \left( t^A \right)^{MN} \,,
\end{equation}
with $C_A$ an element of the $\mathbf{6_{2}}$, and $\left( t^A \right)^{MN}$ the generator corresponding to $\SO{4,4}$ whose only non-zero component is
\begin{equation}
	\left(t^A\right)^{Bz} = \eta^{AB} \,.
\end{equation}
Here $\eta_{MN}$ and $\eta_{AB}$ are the $\SO{4,4}$ and $\SO{3,3}$ invariant metrics, respectively, and are used to raise/lower the corresponding vector indices. Decomposing with respect to the geometric $\SL{3} \times \mathbb{R}^+$ of the $S^3$,  the coordinates on $S^3$ live in the $Y^A = \left( y^i,\, y_i \right)$, $i = 1, 2, 3$, while $C_A = \left( C_i,\, C^i \right)$ naturally contains a 2-form $C^i$. The field strength $H_3 = \partial^A C_A = \partial_i C^i$ is a singlet of $\SO{3,3}$ and thus we see that $\rho^{-1} F_{\ov{M}\ov{N}}{}^{\ov{P}}$ in \eqref{eq:FlatF} is constant and we obtain a consistent truncation with a new 3-form flux. Evaluating the twist matrix
\begin{equation}
	U'_M{}^{\ov{M}} = (e^C)_M{}^N  U_N{}^{\ov{M}} \,,
\end{equation}
explicitly using \eqref{eq:C2S3example} and setting $C^i = \rho^{-1}\,\lambda\, \xi^i$, with $\nabla_i \xi^i = 1$, in the notation of \cite{Eloy:2021fhc}, we obtain precisely the twist matrix used in \cite{Eloy:2021fhc} for the consistent truncation of 6-dimensional ${\cal N}=(1,1)$ supergravity on $S^3$ with $H_3$ flux. %\EM{Give precise form of $C^i$ such that we recover $\lambda$?}

@@

Finally, the metric parametrising this moduli space is:
\begin{equation}
	\begin{aligned}
		\d s^{2}_{\rm mod.} &= -\frac{1}{2}\bigg[e^{2\psi}\left(e^{-\Omega}\,\d\zeta_{1}^{2}+e^{\Omega}\,\d\zeta_{2}^{2}+e^{\ts{\Omega}}\,\d\ts{\zeta}_{1}^{2}+e^{-\ts{\Omega}}\,\d\ts{\zeta}_{2}^{2}\right) +2\,\d\psi^{2} + \frac{1}{2}\,\d\Omega^{2}+ \frac{1}{2}\,\d\ts{\Omega}^{2}\bigg]\,.
	\end{aligned}
\end{equation}
As previously, the deformation breaks the isometry group down to ${\rm U}(1)_{\rm L}\times{\rm U}(1)_{\rm R}\times\ts{{\rm U}(1)}_{\rm L}\times\ts{{\rm U}(1)}_{\rm R}$ and all supersymmetries. Note that for
\begin{equation}
	\zeta_{1}=\pm\zeta_{2},\quad \ts{\zeta}_{1}=\mp\ts{\zeta}_{2},\quad\psi=\Omega=\ts{\Omega}=0,
\end{equation}
the solution~\eqref{eq:metriczeta} reproduces eq.~\eqref{eq:chipsi} for
\begin{equation}
	\chi_{1}=\pm\chi_{2}=\frac{\zeta_{1}}{1+\zeta_{1}^{2}+\ts{\zeta}_{1}^{2}},\quad \ts{\chi}_{1}=\mp\ts{\chi}_{2}=-\frac{\ts{\zeta}_{1}}{1+\zeta_{1}^{2}+\ts{\zeta}_{1}^{2}},\quad e^{2\psi}=\frac{1}{(1+\zeta_{1}^{2}+\ts{\zeta}_{1}^{2})^{2}},
\end{equation}
with ${\cal N}=(4,0)$ (or $(0,4)$) and symmetry enhancement to ${\rm SO}(4)\times{\rm U}(1)^{2}$. There seems to be no other points with (super)-symmetry enhancement in the scalar manifold.

@@

\section{Conclusions} \label{s:Conclusions}
In this paper, we used $\EE$ ExFT to construct the consistent truncation of IIB supergravity around the ${\cal N}=(4,4)$ AdS$_3 \times S^3 \times S^3 \times S^1$ vacuum. Even though the vacuum breaks half of the supersymmetries, we are able to construct a consistent truncation to 3-dimensional maximal gauged supergravity, providing an uplift to the ${\cal N}=16$ gauged supergravity proposed in \cite{Hohm:2005ui} for the 3-dimensional description of this AdS$_3$ vacuum. We constructed our consistent truncation by doing a $S^1$ reduction of the truncation of IIB on $S^3 \times S^3$ \cite{Inverso:2016eet}, and turning on an additional $H_7$ flux on $S^3 \times S^3 \times S^1$, which is necessary to obtain an AdS$_3$ vacuum. We also show that it is not possible to turn on such an additional flux for the other ``dyonic'' consistent truncations on $S^6$, $S^5 \times S^1$, $S^4 \times S^2$ \cite{Guarino:2015jca,Inverso:2016eet} reduced on $S^1$, whilst maintaining a consistent truncation to ${\cal N}=16$ gauged supergravity.
